---
title: "ceRNAR: an R package for identification and analysis of ceRNAs-miRNA triplets" 
author: "\\
    
    Yi-Wen Hsiao (<d08849010@ntu.edu.tw>) and Lin Wang (<linwang071993@gmail.com>)\\
    
    School of Engineering, Dali University"
date: '`r Sys.Date()`'
output:
    BiocStyle::html_document:
      toc: yes
    BiocStyle::pdf_document:
      toc: yes
vignette: >
    %\VignetteIndexEntry{Identification and analysis of miRNA sponge regulation} 
    %\VignettePackage{miRspongeR} 
    % \VignetteEngine{knitr::rmarkdown} 
    % \usepackage[utf8]{inputenc} 
    % \VignetteEncoding{UTF-8}
---

```{r style, echo=FALSE, results="asis", message=FALSE}
BiocStyle::markdown()
knitr::opts_chunk$set(tidy = FALSE,
    warning = FALSE,
    message = FALSE)
```

```{r echo=FALSE, results='hide', message=FALSE}
library(ceRNAR)
```


# Introduction

We present a novel method for identification of ceRNA-miRNA triplets at specific miRNA expression level. Building on rank- and correlation-based approach, we present a computationally
efficient approach to be applied following identification of ceRNA-miRNA triplets that is related to the etiology of diseases.

Our approach leverages
the genom-wide expression profiles and the correlation-based approach to identify ceRNA-miRNA triplets.It is particularly well suited for the studies having both mRNA and miRNA gene expression data. When applied to such studies, ceRNAR provides a scheme for unveiling the novel ceRNAs that regulate the biological systems and analyzing these novel findings for further biological interpretation.

This package also
provides several downstream analyses, including functional, network, survival and peak location analyses, by which to further investigate the identified ceRNA-miRNA triplets and visualize the analytical results to aid in the understanding of the role of such triplets in biological mechanism.


